---
title: "Untitled"
author: "Amelia Ritger"
date: "6/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warnings = FALSE)
```

##############################
# import packages
##############################
```{r}
library(tidyverse) # data wrangling & vizn
library(here) # for reproducibility!
library(chron) # working with times/dates
library(lubridate) # working with times/dates
library(naniar) # working with missing data
library(ggridges) # makes pretty geom_ridge plots

#For ts stuff:
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
```

# metadata - this is a simple csv I built with the file names and associated information; it'll make importing/wrangling all the files in one step a bit easier in Part 1
```{r}
metadata <- read_csv(here("bottom-temp", "csv", "lter-file-names.csv"))
```

# this function imports each of the 5 files, cleans up the data and keeps only the required variables (there are way too many columns to deal with otherwise)
```{r}
import_temp_files <- function(file_name_csv, site){ #file_name_csv and site are generic terms
  read_csv(here::here("bottom-temp", "csv", file_name_csv), col_names = TRUE, col_types = cols(.default = col_character())) %>% 
  select(year, month, day, decimal_time, Temp_bot, Temp_top, Temp_mid) %>% 
  unite(date, year, month, day, sep = "-", remove = FALSE) %>% 
  mutate(time = times(as.numeric(decimal_time))) %>% 
  unite(date_time, date, time, sep = " ") %>% 
  select(date_time, year, month, day, Temp_bot, Temp_top, Temp_mid) %>% 
  mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
         year = as.factor(year),
         month = as.factor(month),
         day = as.numeric(day),
         Temp_top = as.numeric(Temp_top), 
         Temp_mid = as.numeric(Temp_mid), 
         Temp_bot = as.numeric(Temp_bot),
         site = rep(site)) %>% 
  replace_with_na(replace = list(Temp_bot = 9999)) %>% 
  replace_with_na(replace = list(Temp_top = 9999)) %>% 
  replace_with_na(replace = list(Temp_mid = 9999))
}
```

# initialize empty df to fill with cleaned data
```{r}
temp_data <- data.frame(date_time = as.Date(character()),
                        year = factor(),
                        month = factor(),
                        day = numeric(),
                        site = as.character(),
                        Temp_bot = numeric(),
                        Temp_top = numeric(), 
                        Temp_mid = numeric())
```

# for loop reads in/wrangles all files using the import_temp_files() function above
```{r}
for (row in 1:nrow(metadata)) {
  data_row <- metadata[row,]
  file_name_csv <- paste(as.character(data_row[, 1]),".csv", sep="")
  print(file_name_csv) #self check step
  site <- as.character(data_row[,2])
  print(site)
  table <- import_temp_files(file_name_csv, site) 
  temp_data <- rbind(temp_data, table)
}
```

# Save this combinded df as a csv file to work with in other scripts/to avoid having to run the for loop over and over again
```{r}
write.csv(temp_data, here("bottom-temp", "amelia", "all_bottom_temps_2005_2020.csv"), row.names = FALSE)
```

## Sam's plots
```{r}
ggplot(temp_data, aes(x = date_time, y = Temp_bot)) +
  geom_line() +
  #geom_hline(yintercept = 11, linetype = "dashed", color = "red") +
  #geom_hline(yintercept = 16, linetype = "dashed", color = "red") +
  #geom_hline(yintercept = 21, linetype = "dashed", color = "red") +
  facet_wrap(~site, ncol = 1) +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        plot.title = element_text(size = 20, face = "bold"),
        panel.border = element_rect(colour = "black", fil = NA, size = 0.7),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 12),
        legend.position = "none") 
```

```{r}
##############################
# filter out years that don't have data for some sites, add month abbreviations and set factor levels
##############################

cleaned_temp_data <- temp_data %>% 
  #filter(!year %in% c("1999", "2000", "2001", "2002", "2003", "2004")) %>% # remove years that we don't have good data for
  mutate(month = as.factor(month), # there's prob a shortcut for this but I was too lazy to figure it out
         month_name = case_when(
            month == "1" ~ "Jan",
            month == "2" ~ "Feb",
            month == "3" ~ "Mar",
            month == "4" ~ "Apr",
            month == "5" ~ "May",
            month == "6" ~ "Jun",
            month == "7" ~ "Jul",
            month == "8" ~ "Aug",
            month == "9" ~ "Sep",
            month == "10" ~ "Oct",
            month == "11" ~ "Nov",
            month == "12" ~ "Dec"
        )
      ) %>% 
  mutate(month_name = fct_relevel(month_name, "Dec", "Nov", "Oct", "Sep", "Aug", # get months in correct order
                                  "Jul", "Jun", "May", "Apr", "Mar", "Feb", "Jan")) %>% 
  group_by(month_name)

##############################
# identify all unique sites to plot
##############################

unique_sites <- unique(temp_data$site)

##############################
# plot density ridges for all sites
##############################

# pdf(paste("lter_bottom_temp_densityridges_plot", Sys.Date(),".pdf", sep = "_"), onefile = TRUE, width = 10, height = 8)

nap_data <- cleaned_temp_data %>% 
  filter(site=="ARQ")

ggplot(nap_data, aes(x = Temp_bot, y = month_name)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  geom_vline(xintercept = 11, linetype = "dashed", color = "darkgray") +
  geom_vline(xintercept = 16, linetype = "dashed", color = "darkgray") +
  geom_vline(xintercept = 21, linetype = "dashed", color = "darkgray") +
  scale_fill_gradientn(colors = c("steelblue4", "lightblue", "lightslategray", "lightcoral", "indianred4"), name = "Temperature (째C)") +
  scale_x_continuous(breaks = seq(9, 26, by = 2)) +
  labs(x = "Bottom Temperature (째C)", y = "Month", title = site) +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        panel.border = element_rect(colour = "black", fil = NA, size = 0.7),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 12)) 

for(i in 1:length(unique_sites)) {
  
  lter_site <- unique_sites[i]
  
  print(lter_site)
  
  subset <- cleaned_temp_data %>% 
    filter(site == lter_site)
   
  densityridges_plot <- ggplot(subset, aes(x = Temp_bot, y = month_name, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  geom_vline(xintercept = 11, linetype = "dashed", color = "darkgray") +
  geom_vline(xintercept = 16, linetype = "dashed", color = "darkgray") +
  geom_vline(xintercept = 21, linetype = "dashed", color = "darkgray") +
  scale_fill_gradientn(colors = c("steelblue4", "lightblue", "lightslategray", "lightcoral", "indianred4"), name = "Temperature (째C)") +
  scale_x_continuous(breaks = seq(9, 26, by = 2)) +
  labs(x = "Bottom Temperature (째C)", y = "Month", title = lter_site) +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(size = 13),
        panel.border = element_rect(colour = "black", fil = NA, size = 0.7),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(size = 12)) 

  plot(densityridges_plot)
  
}
```

##################################
Amelia's work
##################################

#Get average temperature for each date
```{r}
temp_avg <- temp_data %>% 
  mutate(date = date(date_time),
         mo_yr = yearmonth(date),
         mo = month(mo_yr, label = TRUE)) %>% 
  group_by(date, year, mo, site) %>% 
  summarize(temp_avg = mean(Temp_bot)) %>% 
  ungroup()

#plot average temperature for each month for each site
ggplot(data=temp_avg, aes(x=date, y=temp_avg, color=site)) +
  geom_line() +
  facet_wrap(~site)

ggsave(here("bottom-temp", "amelia", "fig1.png"), height=5, width=8)

#season plot
ggplot(data=temp_avg, aes(x=mo, y=temp_avg, group=year)) +
  geom_line(aes(color=year)) +
  facet_wrap(~site,
             ncol=1,
             scales="free",
             strip.position = "right")
```

Make Naples a timeseries df
```{r}
nap_ts_prep <- cleaned_temp_data %>% 
  group_by(site, year, month, day) %>% 
  summarize(temp = mean(Temp_bot)) %>% #get average temperature for each date
  ungroup() %>%
  unite(date_new, year:day, sep="/") %>% #make a date column in a format to easily run as.Date next
  mutate(date = as.Date(date_new, format="%Y/%m/%d")) %>%  #tell R this is a date
  filter(site=="NAP") %>% 
  drop_na() #drop NA values

#coerce to ts df
nap_ts = ts(data = nap_ts_prep$temp, 
                    start = c(1999,1), 
                    frequency = 365.24)

#summarize
summary(nap_ts)

#visualize
nap_ts %>% plot(xlab = "", las = 1,
                  ylab = expression(Temperature~(degree*C)))
```

From  https://semba-blog.netlify.app/02/22/2019/exploring-time-series-data-in-r/#tab:tab4
ID shifts in mean/variance at Naples
```{r}
library(changepoint)

daily.changepoints = nap_ts %>% 
  changepoint::cpt.meanvar()

daily.changepoints %>% summary()
```

There are two segments with one change point. 

```{r}
library(kableExtra)

daily.changepoints %>% 
  param.est() %>% 
  data.frame() %>% 
  mutate(segment = c("First", "Second")) %>% select(segment, mean, variance) %>%
  kableExtra::kable(format = "html", digits = 2, 
                    caption = "Mean and variance of change point for daily sst",
                    col.names = c("Segment", "Mean", "Variance")) %>%
  kableExtra::column_spec(column = 1:3, width = "3cm") %>%
  kableExtra::add_header_above(c("", "Parameters" = 2))
```

The first segment had a lower mean bottom temperature and lower variance in bottom temperature.

Let's plot the breakpoint
```{r}
daily.changepoints %>% 
  changepoint::plot(xlab = "", ylab = expression(Temperature~(degree*C)), las = 1)
```

Looks like 2010 was the year the change point occurred *for days*.

```{r}
#isolate the exact date for changepoint
#create monthly data set vector from January 1999 to December 2017
analysis.day = seq(dmy(010199), dmy(311217), by = "day")
analysis.day[cpts(daily.changepoints)]
```

December 18, 2010 to be exact

Now, do the same for months (and years)
Month:
```{r}
nap_ts_mo_prep <- cleaned_temp_data %>% 
  group_by(site, year, month) %>% 
  summarize(temp = mean(Temp_bot)) %>% #get average temperature for each date
  ungroup() %>%
  arrange(year, month) %>% 
  filter(site=="NAP") %>% 
  drop_na() #drop NA values

#coerce to ts df
nap_ts_mo = ts(data = nap_ts_mo_prep$temp, 
                    start = c(1999,1), 
                    frequency = 12)

#summarize
summary(nap_ts_mo)

#visualize
nap_ts_mo %>% plot(xlab = "", las = 1,
                  ylab = expression(Temperature~(degree*C)))
```

```{r}
#ID breakpoints
daily.changepoints_mo = nap_ts_mo %>% 
  changepoint::cpt.meanvar()

daily.changepoints_mo %>% summary()

daily.changepoints_mo %>% 
  param.est() %>% 
  data.frame() %>% 
  mutate(segment = c("First", "Second")) %>% select(segment, mean, variance) %>%
  kableExtra::kable(format = "html", digits = 2, 
                    caption = "Mean and variance of change point for daily sst",
                    col.names = c("Segment", "Mean", "Variance")) %>%
  kableExtra::column_spec(column = 1:3, width = "3cm") %>%
  kableExtra::add_header_above(c("", "Parameters" = 2))
```

Two segments, first has lower mean temperature than the second

```{r}
#plot the breakpoint
daily.changepoints_mo %>% 
  changepoint::plot(xlab = "", ylab = expression(Temperature~(degree*C)), las = 1)
#notice that change point is not noticeable, so run it again using a nonparametric function

library(changepoint.np)
daily.changepoints_mo_np = nap_ts_mo %>% changepoint.np::cpt.np() 

daily.changepoints_mo_np %>% changepoint::plot()

#isolate the exact dates for changepoints
#create monthly data set vector from January 1999 to December 2009
analysis.month = seq(dmy(010199), dmy(311209), by = "month")
analysis.month[cpts(daily.changepoints_mo_np)]
```

Change occurred April 2005 and again on November 2006... anything odd about those dates?

Make a Naples heatmap
```{r}
library(oce)

#clean up the data a bit more
day_nap <- cleaned_temp_data %>% 
  group_by(site, year, month, day) %>% 
  summarize(temp = mean(Temp_bot)) %>% 
  ungroup() %>%
  unite(date_new, year:day, sep="/") %>% #make a date column in a format to easily run as.Date next
  mutate(date = as.Date(date_new, format="%Y/%m/%d")) %>%  #tell R this is a date
  select(date, site, temp) %>% 
  filter(site=="NAP", #only look at Naples
         !is.na(temp)) #remove NAs

#define days and years relevant to data
days = 1:366
years = 2001:2019

nap_matrix = matrix(day_nap$temp, nrow = length(days))
imagep(x = years, y = days, z = nap_matrix %>% 
         t(),
       filledContour = F, 
       ylim = c(365,0),
       ylab = "Day", 
       col = oceColors9A(120),
       zlab = expression(Naples-Temperature~(degree *C)))
```

More resources:
Change point detection
https://rpubs.com/richkt/269908

To do: 
- figure out if ts conversion deletes or fills in missing data
- breakpoint analysis for years
- use coefficient of variation to calculate cycle for each site to answer "what is an average day at [site]?"
- compare models and anomalies at each site to assess differences between sites