---
title: "Sea Surface Temperature SBC hourly"
author: "Emily Hardison"
date: "7/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Goal: To search environmental temperature data for intermediate timescales of temperature change.

####Plan: To create a for loop that will:

1. Take an already tidy data set and filter for times of the year where daily variation is >x (user designated range)

2. For all days that have that amount of vairability, calculate average of preceeding 2 weeks

3. calculate average of following 1, 2, 3, 4, 5, 6, 7, 8, 9, ......n days

4. if avg is >preceeding 2 weeks by Y amt, designate it as "use", if not, designate as "toss"

5. filter for usable days after dates with high variability

6. Generate pretty plot of the days

##### Idea from Logan: Sliding Window: Plot the derivative of the data-- to pull out different levels of variability

##### Idea from Krista Auto-regresive models, and thinking about the freshwater ecology candidate -- 

###### From Sam -- simple packages and code to see how related different points are throughout time

###### --> ESM Lab 5

```{r cars, message = FALSE, echo = FALSE, warning = FALSE}
# Attaching Packages
library(tidyverse)
library(janitor)
library(lubridate)
library(ggridges)
```

### Tidying Data

```{r pressure, echo=FALSE}
# Reading in CSV file
sst <- read_csv(here::here("sea-surface-temp", "NOAA_buoy_SBC_hourly_abbrv.csv")) %>% 
  clean_names() %>% # cleaning all names using Janitor
  mutate(
    date_time = make_datetime(year, month, day, hour)) # generating a date.time column using lubridate package
```

### Preliminary Plot of Data
```{r, warning = FALSE, message = FALSE}
# First pas plot to look at SST across years
ggplot(sst, aes(x = water_temp, y = as.factor(month)))+
  geom_density_ridges()+ # From the ggridges package
  ylab("Year") + 
  xlab("Temperature (°C)")+
  theme_bw()


```

### Trying to make a for loop

```{r}
# Generate a new df and calculate the daily max amount of temp variation (daily max-min value for every day in the dataset)
sst_variation <- sst %>% 
  group_by(date(date_time)) %>% 
  summarize(
    daily_max = max(water_temp),
    daily_min = min(water_temp),
    daily_fluc = daily_max - daily_min
  )

# Creating Data Frame with dates where there was High Daily Variation-- >4 degrees in the day
sst_high_var <- sst_variation %>% 
  filter(
    daily_fluc >= 4
  )
# 
# # Need to generate For loop now....
# 
# d <- "1994-06-03"
# for(row in 1:nrow(sst_high_var)){
#   
# # First Step: Calculating the average from the preceeding 2 weeks of data
#   d <- `date(date_time)`
#   
#     previous_avg <- sst %>% 
#     filter(date(date_time) <= date(d) & date(date_time) >= date(d) - 14) %>% 
#     summarize(
#       previous_average = mean(water_temp, na.rm = T)
#     )
#   
# # Second Step: Determining the average of the following XX number of days
#     post_avg <- sst %>% 
#     filter(date(date_time) >= date(d) & date(date_time) <= date(d) + 14) %>% 
#     group_by(date(date_time)) %>% 
#     summarize(
#       daily_max = max(water_temp, na.rm = T),
#       daily_min = min(water_temp, na.rm = T),
#       daily_fluc = daily_max - daily_min, 
#       daily_average = mean(water_temp, na.rm = T)
#     )
  
  # Binding 
    
    
    ### From Sam-- you can make a blank dataframe and then rbind new dataframes to that dataframe
    
    ### naniar package for dealing with Missing data-- different ways to look at it and what to do
    
#}




```









### Parking Lot
```{r}
# First pas plot to look at daily variation across years
# ggplot(sst_variation, aes(x = daily_fluc, y = as.factor(month(`date(date_time)`))))+
#   geom_density_ridges()+ # From the ggridges package
#   ylab("Year") + 
#   xlab("Temperature (°C)")+
#   xlim(0,8)+
#   theme_bw()

```

