---
title: "Sea Surface Temperature SBC hourly"
author: "Emily Hardison"
date: "7/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Original Goal: To search environmental temperature data for intermediate timescales of temperature change.

####Original thoughts for a Plan of action: To create a for loop that will:

1. Take an already tidy data set and filter for times of the year where daily variation is >x (user designated range)

2. For all days that have that amount of vairability, calculate average of preceeding 2 weeks

3. calculate average of following 1, 2, 3, 4, 5, 6, 7, 8, 9, ......n days

4. if avg is >preceeding 2 weeks by Y amt, designate it as "use", if not, designate as "toss"

5. filter for usable days after dates with high variability

6. Generate pretty plot of the days

##### Idea from Logan: Sliding Window: Plot the derivative of the data-- to pull out different levels of variability

##### Idea from Krista Auto-regresive models, and thinking about the freshwater ecology candidate -- 

###### From Sam -- simple packages and code to see how related different points are throughout time

###### --> ESM Lab 5

### What I actually as able to do -- actually just graph the data (mean, min, max) and then look at daily variability throughout the year (daily fluctuations in temp) to start exploring times of year with different amounts of variability

```{r cars, message = FALSE, echo = FALSE, warning = FALSE}
# Attaching Packages
library(tidyverse)
library(janitor)
library(lubridate)
library(ggridges)
library(naniar)
```

### Tidying Data

```{r pressure, echo=FALSE}
# Reading in CSV file
sst <- read_csv(here::here("sea-surface-temp", "NOAA_buoy_SBC_hourly_abbrv.csv")) %>% 
  clean_names() %>% # cleaning all names using Janitor
  mutate(
    year1 = year,
    month1 = month,
    day1 = day,
    hour1 = hour,
    date_time = make_datetime(year, month, day, hour),
    date = make_date(year, month, day)) # generating a date.time column using lubridate package
  
```

### Preliminary Plot of Data
```{r, warning = FALSE, message = FALSE}
# First pas plot to look at SST across years
ggplot(sst, aes(x = water_temp, y = as.factor(month)))+
  geom_density_ridges()+ # From the ggridges package
  ylab("Year") + 
  xlab("Daily Fluctations (°C)")+
  theme_bw()


```



### Making nicer plots of 1) Daily Temperature and 2) Daily Variation

```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.width=6, fig.height=6, fig.align='center'}

summary_sst <- sst %>% 
  group_by(date) %>% 
  summarize(
    max_daily_temp = max(water_temp, na.rm =T),
    avg_daily_temp = mean(water_temp, na.rm =T),
    min_daily_temp = min(water_temp, na.rm =T),
    daily_fluctuation = max_daily_temp - min_daily_temp, na.rm =T) %>% 
  mutate(
    day = case_when(
      month(date) == 1 ~ paste(day(date)),
      month(date) == 2 ~ paste(day(date)+31),
      month(date) == 3 ~ paste(day(date)+59),# Not ACCOUNTING FOR LEAB YEARS :/ 
      month(date) == 4 ~ paste(day(date)+90),#90
      month(date) == 5 ~ paste(day(date)+120),#120
      month(date) == 6 ~ paste(day(date)+151),#151
      month(date) == 7 ~ paste(day(date)+181),#181
      month(date) == 8 ~ paste(day(date)+212),#
      month(date) == 9 ~ paste(day(date)+243),#
      month(date) == 10 ~ paste(day(date)+273),
      month(date) == 11 ~ paste(day(date)+304),#
      month(date) == 12 ~ paste(day(date)+334)
    )) %>% 
  filter(
    max_daily_temp !="-Inf", # removing NA Values
    max_daily_temp !="Inf",
    min_daily_temp !="-Inf",
    max_daily_temp !="Inf",
    avg_daily_temp !="-Inf",
    avg_daily_temp !="Inf"
  )


# There is a lot of missing Data from years 1997, 2014-2018


ggplot(summary_sst, aes(x = as.numeric(day), y = avg_daily_temp,
                        group = year(date), 
                        color = year(date)))+
  geom_line()+ # yearly lines
  ylab("Temperature (°C)") + 
  xlab("Day")+
  theme_bw()


# Plot Smoothing across years (note that some years do not have full data sets and that leap years have not been accounted for -- all years usnig 365 days) 

ggplot(summary_sst)+
  geom_point(aes(x = as.numeric(day), y = avg_daily_temp, na.rm =T), color = "black", alpha = 0.05)+
  geom_point(aes(x = as.numeric(day), y = max_daily_temp, na.rm =T), color = "red", alpha = 0.05)+
  geom_point(aes(x = as.numeric(day), y = min_daily_temp, na.rm =T), color = "blue", alpha = 0.05)+
  geom_smooth(aes(x = as.numeric(day), y = avg_daily_temp, na.rm =T), color = "black", size = 1.5)+ # yearly lines
  geom_smooth(aes(x = as.numeric(day), y = max_daily_temp, na.rm =T), color = "red", size = 1.5)+
  geom_smooth(aes(x = as.numeric(day), y = min_daily_temp, na.rm =T), color = "blue", size = 1.5)+
  ylab("Temperature (°C)") + 
  xlab("Day")+
  scale_y_continuous(limits = c(8, 24), expand = c(0,0), breaks = c(10, 14, 18, 22))+ # Y-axis
  geom_hline(yintercept = c(10, 14, 18, 22), linetype = "dashed", alpha = 0.5)+# adding hlines
  theme_classic()+
  theme(axis.text = element_text(size=12, color = "black"), 
        axis.title.y= element_text(size=15,  color = "black"), 
        axis.title.x= element_text(size=15), 
        legend.position = "none", 
        axis.line = element_line(colour = 'black', size = 0.75), 
        plot.title = element_text(hjust = 0.5, size=20))

# Plot Smoothing across years without points(note that some years do not have full data sets and that leap years have not been accounted for -- all years usnig 365 days) 

ggplot(summary_sst)+
  #geom_point(aes(x = as.numeric(day), y = avg_daily_temp), color = "black", alpha = 0.1)+
  #geom_point(aes(x = as.numeric(day), y = max_daily_temp), color = "red", alpha = 0.1)+
  #geom_point(aes(x = as.numeric(day), y = min_daily_temp), color = "blue", alpha = 0.1)+
  geom_smooth(aes(x = as.numeric(day), y = avg_daily_temp), color = "black", size = 1.5)+ # yearly lines
  geom_smooth(aes(x = as.numeric(day), y = max_daily_temp), color = "red", size = 1.5)+
  geom_smooth(aes(x = as.numeric(day), y = min_daily_temp), color = "blue", size = 1.5)+
  ylab("Temperature (°C)") + 
  xlab("Day")+
  scale_y_continuous(limits = c(8, 24), expand = c(0,0), breaks = c(12, 16, 20, 24))+ # Y-axis
  geom_hline(yintercept = c(12, 16, 20), linetype = "dashed", alpha = 0.5)+# adding hlines
  theme_classic()+
  theme(axis.text = element_text(size=12, color = "black"), 
        axis.title.y= element_text(size=15,  color = "black"), 
        axis.title.x= element_text(size=15), 
        legend.position = "none", 
        axis.line = element_line(colour = 'black', size = 0.75), 
        plot.title = element_text(hjust = 0.5, size=20))



# Daily Fluctuations in temperature 

ggplot(summary_sst, aes(x = as.numeric(day), y = daily_fluctuation))+
  geom_point(aes(group = year(date), color = year(date)), alpha = 0.25)+
  geom_smooth(color = "black")+ # yearly lines
  ylab("Temperature (°C)") + 
  xlab("Day")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2, 4, 6, 8, 10))+ # Y-axis
  #geom_hline(yintercept = c(10, 13, 16, 19), linetype = "dashed", alpha = 0.5)+# adding hlines
  theme_classic()+
  theme(axis.text.y = element_text(size=12, color = "black"), 
        axis.text.x = element_blank(), 
        axis.title.y= element_text(size=15,  color = "black"), 
        axis.title.x= element_text(size=15), 
        legend.position = "none", 
        axis.line = element_line(colour = 'black', size = 0.75), 
        plot.title = element_text(hjust = 0.5, size=20))+
  facet_wrap(~year(date))



```




### How to deal with missing data? Want to discuss this more in the group. Sam had sent around this resource from Allison Horst:
https://allisonhorst.shinyapps.io/missingexplorer/#section-introduction

There are quite a lot of missing years in the data set. The naniar package can tell me the proportion of missing data relative to the entire data set

```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.width=6, fig.height=6, fig.align='center'}
miss_var_summary(sst) # naniar function that tells you the proportion of missing data for each variable
```


There is a lot of missing data for entire years, and I am not sure how to deal with this, especially since most of the missing data is in the more recent years.

Questions for group:
1) How to graph variability? Do you like the idea of doing max daily var across the year?
2) What to do when the recent data is missing? I.e. if you had an experiment in 2020, but there are large holes in the data after 2016? 

To make a daily variability plot, I will just take the latest full year of data (2016) and make a nicer version of that plot


```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.width=6, fig.height=6, fig.align='center'}


summary_sst_2016 <- summary_sst %>% 
  filter(year(date) == 2017)

# Daily Fluctuations in temperature 

ggplot(summary_sst_2016, aes(x = as.numeric(day), y = daily_fluctuation))+
  geom_point()+
  geom_smooth(color = "black")+ # yearly lines
  ylab("Temperature (°C)") + 
  xlab("Day")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2, 4, 6, 8, 10))+ # Y-axis
  #geom_hline(yintercept = c(10, 13, 16, 19), linetype = "dashed", alpha = 0.5)+# adding hlines
  theme_classic()+
  theme(axis.text = element_text(size=12, color = "black"), 
        axis.title.y= element_text(size=15,  color = "black"), 
        axis.title.x= element_text(size=15), 
        legend.position = "none", 
        axis.line = element_line(colour = 'black', size = 0.75), 
        plot.title = element_text(hjust = 0.5, size=20))

```


I'm interested in how variability from 1994-1996 compares to 2015-2017. Going to filter the data for a few years early in the data set to visualize how to "smoothed" data compare to the more recent years. Not feeling super confident in this approach, but thought it could be cool to look at.

```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.width=6, fig.height=6, fig.align='center'}
summary_sst_compare<- summary_sst %>% 
  filter(year(date) %in% c(1994, 1995, 1996, 2015, 2016, 2017)) %>% 
  mutate(time = case_when(
    year(date) %in% c(1994, 1995, 1996) ~ "90's",
    year(date) %in% c(2015, 2016, 2017) ~ "recent"))

ggplot(summary_sst_compare, aes(x = as.numeric(day), y = daily_fluctuation, color = time, group = year(date)))+
  #geom_point()+
  geom_smooth(alpha = 0.1)+ # yearly lines
  labs(y = "Temperature (°C)",
       x="Day",
       color = "")+
  scale_y_continuous(limits = c(0, 8), expand = c(0,0), breaks = c(2, 4, 6, 8))+ # Y-axis
  #geom_hline(yintercept = c(10, 13, 16, 19), linetype = "dashed", alpha = 0.5)+# adding hlines
  theme_classic()+
  scale_color_manual(values = c("blue", "orange"))+
  theme(axis.text = element_text(size=12, color = "black"), 
        axis.title.y= element_text(size=15,  color = "black"), 
        axis.title.x= element_text(size=15), 
        legend.position = "bottom", 
        axis.line = element_line(colour = 'black', size = 0.75), 
        plot.title = element_text(hjust = 0.5, size=20))

```






### Trying to make a for loop (and failing)

```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.width=6, fig.height=6, fig.align='center'}
# Generate a new df and calculate the daily max amount of temp variation (daily max-min value for every day in the dataset)
sst_variation <- sst %>% 
  group_by(date(date_time)) %>% 
  summarize(
    daily_max = max(water_temp),
    daily_min = min(water_temp),
    daily_fluc = daily_max - daily_min
  )

# Creating Data Frame with dates where there was High Daily Variation-- >4 degrees in the day
sst_high_var <- sst_variation %>% 
  filter(
    daily_fluc >= 4
  )
# 
# # Need to generate For loop now....
# 
# d <- "1994-06-03"
# for(row in 1:nrow(sst_high_var)){
#   
# # First Step: Calculating the average from the preceeding 2 weeks of data
#   d <- `date(date_time)`
#   
#     previous_avg <- sst %>% 
#     filter(date(date_time) <= date(d) & date(date_time) >= date(d) - 14) %>% 
#     summarize(
#       previous_average = mean(water_temp, na.rm = T)
#     )
#   
# # Second Step: Determining the average of the following XX number of days
#     post_avg <- sst %>% 
#     filter(date(date_time) >= date(d) & date(date_time) <= date(d) + 14) %>% 
#     group_by(date(date_time)) %>% 
#     summarize(
#       daily_max = max(water_temp, na.rm = T),
#       daily_min = min(water_temp, na.rm = T),
#       daily_fluc = daily_max - daily_min, 
#       daily_average = mean(water_temp, na.rm = T)
#     )
  
  # Binding 
    
    
    ### From Sam-- you can make a blank dataframe and then rbind new dataframes to that dataframe
    
    ### naniar package for dealing with Missing data-- different ways to look at it and what to do
    
#}




```






### Parking Lot
```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.width=6, fig.height=6, fig.align='center'}
# First pas plot to look at daily variation across years
# ggplot(sst_variation, aes(x = daily_fluc, y = as.factor(month(`date(date_time)`))))+
#   geom_density_ridges()+ # From the ggridges package
#   ylab("Year") + 
#   xlab("Temperature (°C)")+
#   xlim(0,8)+
#   theme_bw()

```

