---
title: "Sea Surface Temperature SBC hourly"
author: "Emily Hardison"
date: "7/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Goal: To search environmental temperature data for intermediate timescales of temperature change.

####Plan: To create a for loop that will:

1. Take an already tidy data set and filter for times of the year where daily variation is >x (user designated range)

2. For all days that have that amount of vairability, calculate average of preceeding 2 weeks

3. calculate average of following 1, 2, 3, 4, 5, 6, 7, 8, 9, ......n days

4. if avg is >preceeding 2 weeks by Y amt, designate it as "use", if not, designate as "toss"

5. filter for usable days after dates with high variability

6. Generate pretty plot of the days

##### Idea from Logan: Sliding Window: Plot the derivative of the data-- to pull out different levels of variability

##### Idea from Krista Auto-regresive models, and thinking about the freshwater ecology candidate -- 

###### From Sam -- simple packages and code to see how related different points are throughout time

###### --> ESM Lab 5

```{r cars, message = FALSE, echo = FALSE, warning = FALSE}
# Attaching Packages
library(tidyverse)
library(janitor)
library(lubridate)
library(ggridges)
```

### Tidying Data

```{r pressure, echo=FALSE}
# Reading in CSV file
sst <- read_csv(here::here("sea-surface-temp", "NOAA_buoy_SBC_hourly_abbrv.csv")) %>% 
  clean_names() %>% # cleaning all names using Janitor
  mutate(
    year1 = year,
    month1 = month,
    day1 = day,
    hour1 = hour,
    date_time = make_datetime(year, month, day, hour),
    date = make_date(year, month, day)) # generating a date.time column using lubridate package
  
```

### Preliminary Plot of Data
```{r, warning = FALSE, message = FALSE}
# First pas plot to look at SST across years
ggplot(sst, aes(x = water_temp, y = as.factor(month)))+
  geom_density_ridges()+ # From the ggridges package
  ylab("Year") + 
  xlab("Daily Fluctations (°C)")+
  theme_bw()


```



### Making nicer plots of 1) Daily Temperature and 2) Daily Variation

```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.width=6, fig.height=6}

summary_sst <- sst %>% 
  group_by(date) %>% 
  summarize(
    max_daily_temp = max(water_temp),
    avg_daily_temp = mean(water_temp),
    min_daily_temp = min(water_temp),
    daily_fluctuation = max_daily_temp - min_daily_temp) %>% 
  mutate(
    day = case_when(
      month(date) == 1 ~ paste(day(date)),
      month(date) == 2 ~ paste(day(date)+31),
      month(date) == 3 ~ paste(day(date)+59),# Not ACCOUNTING FOR LEAB YEARS :/ 
      month(date) == 4 ~ paste(day(date)+90),#90
      month(date) == 5 ~ paste(day(date)+120),#120
      month(date) == 6 ~ paste(day(date)+151),#151
      month(date) == 7 ~ paste(day(date)+181),#181
      month(date) == 8 ~ paste(day(date)+212),#
      month(date) == 9 ~ paste(day(date)+243),#
      month(date) == 10 ~ paste(day(date)+273),
      month(date) == 11 ~ paste(day(date)+304),#
      month(date) == 12 ~ paste(day(date)+334)
    ))


# There is a lot of missing Data from years 1997, 2014-2018


ggplot(summary_sst, aes(x = as.numeric(day), y = avg_daily_temp,
                        group = year(date), 
                        color = year(date)))+
  geom_line()+ # yearly lines
  ylab("Temperature (°C)") + 
  xlab("Day")+
  theme_bw()


# Plot Smoothing across years (note that some years do not have full data sets and that leap years have not been accounted for -- all years usnig 365 days) 

ggplot(summary_sst)+
  geom_point(aes(x = as.numeric(day), y = avg_daily_temp), color = "black", alpha = 0.1)+
  geom_point(aes(x = as.numeric(day), y = max_daily_temp), color = "red", alpha = 0.1)+
  geom_point(aes(x = as.numeric(day), y = min_daily_temp), color = "blue", alpha = 0.1)+
  geom_smooth(aes(x = as.numeric(day), y = avg_daily_temp), color = "black", size = 1.5)+ # yearly lines
  geom_smooth(aes(x = as.numeric(day), y = max_daily_temp), color = "red", size = 1.5)+
  geom_smooth(aes(x = as.numeric(day), y = min_daily_temp), color = "blue", size = 1.5)+
  ylab("Temperature (°C)") + 
  xlab("Day")+
  scale_y_continuous(limits = c(8, 24), expand = c(0,0), breaks = c(10, 14, 18, 22))+ # Y-axis
  geom_hline(yintercept = c(10, 14, 18, 22), linetype = "dashed", alpha = 0.5)+# adding hlines
  theme_classic()+
  theme(axis.text = element_text(size=12, color = "black"), 
        axis.title.y= element_text(size=15,  color = "black"), 
        axis.title.x= element_text(size=15), 
        legend.position = "none", 
        axis.line = element_line(colour = 'black', size = 0.75), 
        plot.title = element_text(hjust = 0.5, size=20))

# Plot Smoothing across years without points(note that some years do not have full data sets and that leap years have not been accounted for -- all years usnig 365 days) 

ggplot(summary_sst)+
  #geom_point(aes(x = as.numeric(day), y = avg_daily_temp), color = "black", alpha = 0.1)+
  #geom_point(aes(x = as.numeric(day), y = max_daily_temp), color = "red", alpha = 0.1)+
  #geom_point(aes(x = as.numeric(day), y = min_daily_temp), color = "blue", alpha = 0.1)+
  geom_smooth(aes(x = as.numeric(day), y = avg_daily_temp), color = "black", size = 1.5)+ # yearly lines
  geom_smooth(aes(x = as.numeric(day), y = max_daily_temp), color = "red", size = 1.5)+
  geom_smooth(aes(x = as.numeric(day), y = min_daily_temp), color = "blue", size = 1.5)+
  ylab("Temperature (°C)") + 
  xlab("Day")+
  scale_y_continuous(limits = c(8, 24), expand = c(0,0), breaks = c(12, 16, 20, 24))+ # Y-axis
  geom_hline(yintercept = c(12, 16, 20), linetype = "dashed", alpha = 0.5)+# adding hlines
  theme_classic()+
  theme(axis.text = element_text(size=12, color = "black"), 
        axis.title.y= element_text(size=15,  color = "black"), 
        axis.title.x= element_text(size=15), 
        legend.position = "none", 
        axis.line = element_line(colour = 'black', size = 0.75), 
        plot.title = element_text(hjust = 0.5, size=20))



# Daily Fluctuations in temperature 

ggplot(summary_sst, aes(x = as.numeric(day), y = daily_fluctuation))+
  geom_point(aes(group = year(date), color = year(date)), alpha = 0.25)+
  geom_smooth(color = "black")+ # yearly lines
  ylab("Temperature (°C)") + 
  xlab("Day")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2, 4, 6, 8, 10))+ # Y-axis
  #geom_hline(yintercept = c(10, 13, 16, 19), linetype = "dashed", alpha = 0.5)+# adding hlines
  theme_classic()+
  theme(axis.text = element_text(size=12, color = "black"), 
        axis.title.y= element_text(size=15,  color = "black"), 
        axis.title.x= element_text(size=15), 
        legend.position = "none", 
        axis.line = element_line(colour = 'black', size = 0.75), 
        plot.title = element_text(hjust = 0.5, size=20))+
  facet_wrap(~year(date))



```

### Trying to make a for loop

```{r}
# Generate a new df and calculate the daily max amount of temp variation (daily max-min value for every day in the dataset)
sst_variation <- sst %>% 
  group_by(date(date_time)) %>% 
  summarize(
    daily_max = max(water_temp),
    daily_min = min(water_temp),
    daily_fluc = daily_max - daily_min
  )

# Creating Data Frame with dates where there was High Daily Variation-- >4 degrees in the day
sst_high_var <- sst_variation %>% 
  filter(
    daily_fluc >= 4
  )
# 
# # Need to generate For loop now....
# 
# d <- "1994-06-03"
# for(row in 1:nrow(sst_high_var)){
#   
# # First Step: Calculating the average from the preceeding 2 weeks of data
#   d <- `date(date_time)`
#   
#     previous_avg <- sst %>% 
#     filter(date(date_time) <= date(d) & date(date_time) >= date(d) - 14) %>% 
#     summarize(
#       previous_average = mean(water_temp, na.rm = T)
#     )
#   
# # Second Step: Determining the average of the following XX number of days
#     post_avg <- sst %>% 
#     filter(date(date_time) >= date(d) & date(date_time) <= date(d) + 14) %>% 
#     group_by(date(date_time)) %>% 
#     summarize(
#       daily_max = max(water_temp, na.rm = T),
#       daily_min = min(water_temp, na.rm = T),
#       daily_fluc = daily_max - daily_min, 
#       daily_average = mean(water_temp, na.rm = T)
#     )
  
  # Binding 
    
    
    ### From Sam-- you can make a blank dataframe and then rbind new dataframes to that dataframe
    
    ### naniar package for dealing with Missing data-- different ways to look at it and what to do
    
#}




```









### Parking Lot
```{r}
# First pas plot to look at daily variation across years
# ggplot(sst_variation, aes(x = daily_fluc, y = as.factor(month(`date(date_time)`))))+
#   geom_density_ridges()+ # From the ggridges package
#   ylab("Year") + 
#   xlab("Temperature (°C)")+
#   xlim(0,8)+
#   theme_bw()

```

