---
title: "Bottom Temp Analysis for LTER site ALE"
author: "Terra Dressler"
date: "July 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Packages
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
library(chron)
library(lubridate)
library(naniar)
library(ggsci)

#How to extract color ID #'s from palettes
library(scales)
show_col(pal_material("cyan", n = 10)(10)) 
show_col(pal_rickandmorty("schwifty", alpha = 0.8)(12))

```


Note: I read in temp data using the code on LTER website and wrote a csv file using write_csv


2. Organize
```{r, warning=FALSE, message=FALSE}

#dataframe dt1 read in using the knb-lter-sbc R code from the ALE page on the LTER website
ALE_data <-  read_csv(here::here("bottom-temp", "terra", "ALE_LTER_data.csv"), col_names = TRUE, col_types = cols(.default = col_character())) %>% 
  unite(date, year, month, day, sep = "-", remove = FALSE) %>% 
  mutate(time = times(as.numeric(Decimal.time))) %>% 
  unite(date_time, date, time, sep = " ") %>% 
  select(date_time, year, month, day, Temp_bot, Temp_top, Temp_mid) %>% 
  mutate(date_time = as.POSIXct(date_time, "%Y-%m-%d %H:%M:%S", tz = "GMT"),
         year = as.numeric(year),
         month = as.numeric(month),
         day = as.numeric(day),
         Temp_top = as.numeric(Temp_top), 
         Temp_mid = as.numeric(Temp_mid), 
         Temp_bot = as.numeric(Temp_bot),
         site = rep("ALE")) 
  
  
```

3. Data exploration: Visualize all data
```{r}
ggplot(ALE_data) +
  geom_line(aes(x = date_time, y = Temp_bot), color = "#1967D2FF") +
  theme_classic() +
  labs(x = "Time", y = "Bottom Temperature")
  
```


4. Data exploration: monthly bottom temp patterns

```{r}
ALE_monthly <- ALE_data %>% 
  na.omit() %>% 
  group_by(year, month) %>% 
  summarise(mean_month = mean(Temp_bot), mean_top10 = mean(Temp_bot[Temp_bot>=quantile(Temp_bot, 0.9)]), mean_bottom10 = mean(Temp_bot[Temp_bot<=quantile(Temp_bot, 0.1)]), n = n()) %>%  
  mutate(range_month = mean_top10-mean_bottom10) 

#monthly means
ggplot(ALE_monthly, aes(x = month, y = mean_month, color = year))+
  geom_point(size = 2)+
 scale_color_material(palette = "purple")+
  labs(y = "Mean Monthly Temperature", x = "Month (Jan-Dec)")+
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12))+
  theme_bw()

#monthly max
ggplot(ALE_monthly, aes(x = month, y =  mean_top10, color = year))+
  geom_point(size = 2)+
 scale_color_material(palette = "red")+
  labs(y = "Max Monthly Temperature", x = "Month (Jan-Dec)")+
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12))+
  theme_bw()

#monthly min
ggplot(ALE_monthly, aes(x = month, y = mean_bottom10, color = year))+
  geom_point(size = 2)+
 scale_color_material(palette = "blue")+
  labs(y = "Min Monthly Temperature", x = "Month (Jan-Dec)")+
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12))+
  theme_bw()

#monthly range
ggplot(ALE_monthly, aes(x = month, y = range_month, color = year))+
  geom_point(size = 2)+
 scale_color_material(palette = "cyan")+
  labs(y = "Monthly Temp Range", x = "Month (Jan-Dec)")+
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12))+
  theme_bw()

```

Note: I took the mean top 10% and bottom 10% instead of just the max and min values to accound for potential anomalous results / measurement errors.


Note: I am using a color gradient to see if there are any perceptible effects of time (1999-2019) in these exploratory plots

What I notice: 
1. Coldest monthly temps tend to occur around April
2. Warmest monthly temps tend to occur July-September
3. Temp variability seems to peak around July
4. Highest temps seem to be co-occuring with highest variability 

BUT, this way of visualizing does not capture daily patterns

5. Data exploration: daily bottom temperatures
```{r}
#month_names <- Vectorize(function(M) c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")[M])

ALE_daily <- ALE_data %>% 
  group_by(year, month, day) %>% 
  na.omit() %>% 
  summarise(mean_day = mean(Temp_bot), mean_top10 = mean(Temp_bot[Temp_bot>=quantile(Temp_bot, 0.9)]), mean_bottom10 = mean(Temp_bot[Temp_bot<=quantile(Temp_bot, 0.1)]), n = n()) %>% 
  mutate(range_day = mean_top10-mean_bottom10) %>%
  unite(month_day, month, day, sep = "-") 

ggplot(ALE_daily, aes(x = month_day, y = mean_day, color = year))+
  geom_point(size = 2)+
 scale_color_material(palette = "purple")+
  scale_x_discrete(breaks = c(1-1, 2-1, 3-1, 4-1, 5-1, 6-1, 7-1, 8-1, 9-1, 10-1, 11-1, 12-1))+
  scale_y_continuous(breaks = c(10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22))+
  labs(y = "Mean Temperature", x = "Day (Jan 1 - Dec 31)")+
  theme_bw()

ggplot(ALE_daily, aes(x = month_day, y =  mean_top10, color = year))+
  geom_point(size = 2)+
 scale_color_material(palette = "red")+
  scale_x_discrete(breaks = c(1-1, 2-1, 3-1, 4-1, 5-1, 6-1, 7-1, 8-1, 9-1, 10-1, 11-1, 12-1))+
  scale_y_continuous(breaks = c(10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22))+
  labs(y = "Max Temperature", x = "Day (Jan 1 - Dec 31)")+
  theme_bw()

ggplot(ALE_daily, aes(x = month_day, y = mean_bottom10, color = year))+
  geom_point(size = 2)+
 scale_color_material(palette = "blue")+
scale_x_discrete(breaks = c(1-1, 2-1, 3-1, 4-1, 5-1, 6-1, 7-1, 8-1, 9-1, 10-1, 11-1, 12-1))+
  scale_y_continuous(breaks = c(9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22))+
  labs(y = "Min Temperature", x = "Day (Jan 1 - Dec 31)")+
  theme_bw()

ggplot(ALE_daily, aes(x = month_day, y = range_day, color = year))+
  geom_point(size = 2)+
 scale_color_material(palette = "cyan")+
  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7))+
  scale_x_discrete(breaks = c(1-1, 2-1, 3-1, 4-1, 5-1, 6-1, 7-1, 8-1, 9-1, 10-1, 11-1, 12-1))+
  labs(y = "Temperature Range", x = "Day (Jan 1 - Dec 31)")+
  theme_bw()

  
```


Stratification
